### 2016-02-03 PaperThoughts.Rnw
add some thoughts on paper

\documentclass[a4paper]{report}
\usepackage[margin=.75in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,array}
\usepackage{booktabs}

%% load any required packages here

\usepackage{adjustbox}            %% to align tops of minipages
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{afterpage}

\usepackage{color}
\definecolor{purple}{rgb}{.4,0,.8}
\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\st}[1]{{\color{purple} #1}}

\begin{document}

\LARGE{\bf Thoughts about Network Visual Inference Paper}

\begin{enumerate}
\item Need to complete lineups in 2-3 weeks
\item Lineups: 
  \begin{enumerate}
  \item What sizes? 3, 6, 9, 12
  \item What data? Football, olga's data, teenage friends data?
  \item Other things to consider: coloring of nodes? 
  \end{enumerate}
\item Get simulated networds from models from RSiena
\item Figure out another way to fit a model and then simulate from it? 
\end{enumerate}

Possible section outline for paper:
\begin{enumerate}
\item Introduction - visual inference, networks
\item Data and Model fitting & simulation
\item Lineup methods
\item Results
\item Conclusions and Discussion
\end{enumerate}

\end{document}