### 2016-02-09 paper/network-inference.Rnw

\documentclass[journal]{vgtc}                % final (journal style)
%\documentclass[review,journal]{vgtc}         % review (journal style)
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
%\documentclass[electronic,journal]{vgtc}     % electronic version, journal

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures other than the optional teaser is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures other than the optional teaser on the
%% first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}

%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% This turns references into clickable hyperlinks.
\usepackage[bookmarks,backref=true,linkcolor=black]{hyperref} %,colorlinks
\hypersetup{
  pdfauthor = {},
  pdftitle = {},
  pdfsubject = {},
  pdfkeywords = {},
  colorlinks=true,
  linkcolor= black,
  citecolor= black,
  pageanchor=true,
  urlcolor = black,
  plainpages = false,
  linktocpage
}

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{Visual inference for networks}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Samantha Tyner, Heike Hofmann \textit{Member, IEEE}}
\authorfooter{
%% insert punctuation at end of each item
\item
 Samantha Tyner is with the Department of Statistics, Iowa State University. E-mail: sctyner@iastate.edu.
\item
 Heike Hofmann is with the Department of Statistics, Iowa State University. E-mail: hofmann@iastate.edu.
}

%other entries to be set up for journal
\shortauthortitle{Tyner \MakeLowercase{\textit{et al.}}: Visual inference for networks}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

%% Abstract section.
\abstract{abstract is here} % end of abstract

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Human subject testing, visual inference, Siena models, network model diagnostics}

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CCScat'' command takes four arguments.

% HH: do we need those classifications?
% \CCScatlist{ % not used in journal version
%  \CCScat{K.6.1}{Management of Computing and Information Systems}%
% {Project and People Management}{Life Cycle};
%  \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
% }

%% Uncomment below to include a teaser figure.
  \teaser{
 \centering
 Teaser picture goes here
  \caption{In the Clouds: Vancouver from Cypress Mountain.}
  }

%% Uncomment below to disable the manuscript note
%\renewcommand{\manuscriptnotetxt}{}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

%% \section{Introduction} %for journal use above \firstsection{..} instead

\begin{enumerate}
\item Need to complete lineups in 2-3 weeks
\item Lineups: 
  \begin{enumerate}
  \item What sizes? 3, 6, 9, 12
  \item What data? Football, olga's data, teenage friends data?
  \item Other things to consider: coloring of nodes? 
  \item use replicates: different seeds for the same data produces different layouts -
  different draws from the model
  \end{enumerate}
\item Get simulated networds from models from RSiena
\item Figure out another way to fit a model and then simulate from it? 
\end{enumerate}

Possible section outline for paper:
\begin{enumerate}
\item Introduction - visual inference, networks
\item Data and Model fitting \& simulation
\item Lineup methods
\item Results
\item Conclusions and Discussion
\end{enumerate}

\section{Exposition}

Equations are cool.

\begin{equation}
\sum_{j=1}^{z} j = \frac{z(z+1)}{2}
\end{equation}

Notice the caption at the top of Table~\ref{vis_accept}.

\begin{table}
%% Table captions on top in journal version
 \caption{SciVis Paper Acceptance Rate: 1994-2006}
 \label{vis_accept}
 \scriptsize
 \begin{center}
   \begin{tabular}{cccc}
     Year & Submitted & Accepted & Accepted (\%)\\
   \hline
     2004 & 167 & 46 & 27.6\\
     2005 & 268 & 88 & 32.8\\
     2006 & 228 & 63 & 27.6
   \end{tabular}
 \end{center}
\end{table}

\subsection{Mezcal Head}

Citing happens only with cite - like this for the lineup protocol~\cite{buja:2009}, and for RSiena~\cite{RSiena}. Footnotes are ok%
\footnote{Footnotes appear at the bottom of the column}.

\section{Conclusion}

%% if specified like this the section will be committed in review mode
\acknowledgments{
The authors wish to thank A, B, C. This work was supported in part by
a grant from XYZ.}

\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{references}
\end{document}