### 2016-02-10 paper/network-inference.Rnw
rearrange folders & put code in paper

\documentclass[journal]{vgtc}                % final (journal style)
%\documentclass[review,journal]{vgtc}         % review (journal style)
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
%\documentclass[electronic,journal]{vgtc}     % electronic version, journal

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures other than the optional teaser is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures other than the optional teaser on the
%% first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{color}
\definecolor{purple}{rgb}{.4,0,.8}
\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\st}[1]{{\color{purple} #1}}

%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% This turns references into clickable hyperlinks.
\usepackage[bookmarks,backref=true,linkcolor=black]{hyperref} %,colorlinks
\hypersetup{
  pdfauthor = {},
  pdftitle = {},
  pdfsubject = {},
  pdfkeywords = {},
  colorlinks=true,
  linkcolor= black,
  citecolor= black,
  pageanchor=true,
  urlcolor = black,
  plainpages = false,
  linktocpage
}

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{Visual inference for networks}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Samantha Tyner, Heike Hofmann \textit{Member, IEEE}}
\authorfooter{
%% insert punctuation at end of each item
\item
 Samantha Tyner is with the Department of Statistics, Iowa State University. E-mail: sctyner@iastate.edu.
\item
 Heike Hofmann is with the Department of Statistics, Iowa State University. E-mail: hofmann@iastate.edu.
}

%other entries to be set up for journal
\shortauthortitle{Tyner \MakeLowercase{\textit{et al.}}: Visual inference for networks}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

%% Abstract section.
\abstract{abstract is here} % end of abstract

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Human subject testing, visual inference, Siena models, network model diagnostics}

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CCScat'' command takes four arguments.

% HH: do we need those classifications?
% \CCScatlist{ % not used in journal version
%  \CCScat{K.6.1}{Management of Computing and Information Systems}%
% {Project and People Management}{Life Cycle};
%  \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
% }

%% Uncomment below to include a teaser figure.
  \teaser{
 \centering
 Teaser picture goes here
  \caption{In the Clouds: Vancouver from Cypress Mountain.}
  }

%% Uncomment below to disable the manuscript note
%\renewcommand{\manuscriptnotetxt}{}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

<<packages, echo=FALSE, warning=FALSE, message=FALSE>>=
library(RSiena)
library(network)
library(ggplot2)
library(geomnet)
@
%% \section{Introduction} %for journal use above \firstsection{..} instead

\begin{enumerate}
\item Need to complete lineups in 2-3 weeks
\item Lineups: 
  \begin{enumerate}
  \item What sizes? 3, 6, 9, 12
  \item What data? Football, olga's data, teenage friends data?
  \item Other things to consider: coloring of nodes? 
  \item use replicates: different seeds for the same data produces different layouts -
  different draws from the model
  \end{enumerate}
\item Get simulated networds from models from RSiena
\item Figure out another way to fit a model and then simulate from it? 
\end{enumerate}

Possible section outline for paper:
\begin{enumerate}
\item Introduction - visual inference, networks
\item Data and Model fitting \& simulation
\item Lineup methods
\item Results
\item Conclusions and Discussion
\end{enumerate}

\section{Examples \st{??}}

\st{
One of the data sets provided with RSiena is a subset of 50 female respondants from the \emph{Teenage Friends and Lifestyle Study} XXX citation??? XXX. There are three friendship networks provided at three timepoints representing the relationships between the 50 teenage girls in the data. There are also actor-level variables observed at each time period, such as a girl's drinking and smoking behavior. These networks are shown in \autoref{fig:friendship_waves}
XX autoref not working?? XX
<<friendship_waves, echo=FALSE, message=FALSE, warning=FALSE>>=
friend.data.w1 <- as.matrix(read.table("../Data/s50_data/s50-network1.dat"))
friend.data.w2 <- as.matrix(read.table("../Data/s50_data/s50-network2.dat"))
friend.data.w3 <- as.matrix(read.table("../Data/s50_data/s50-network3.dat"))
drink <- as.matrix(read.table("../Data/s50_data/s50-alcohol.dat"))

friend.data.edges <- data.frame(as.edgelist(as.network(friend.data.w1)))
friend.data.nodes <- data.frame(ids = 1:50, drink = drink[,1])
friend.data.net1 <- merge(friend.data.edges, friend.data.nodes, by.x="X1", by.y = "ids", all = T)
friend.data.edges <- data.frame(as.edgelist(as.network(friend.data.w2)))
friend.data.nodes <- data.frame(ids = 1:50, drink = drink[,2])
friend.data.net2 <- merge(friend.data.edges, friend.data.nodes, by.x="X1", by.y = "ids", all = T)
friend.data.edges <- data.frame(as.edgelist(as.network(friend.data.w3)))
friend.data.nodes <- data.frame(ids = 1:50, drink = drink[,3])
friend.data.net3 <- merge(friend.data.edges, friend.data.nodes, by.x="X1", by.y = "ids", all = T)

friend.data.net.all <- rbind(friend.data.net1, friend.data.net2, friend.data.net3)
names(friend.data.net.all)[1:2] <- c('from', 'to')
friend.data.net.all$period <- rep(1:3, c(nrow(friend.data.net1),
                                  nrow(friend.data.net2),
                                  nrow(friend.data.net3)))

ggplot(data=friend.data.net.all, aes(from_id = from, to_id = to)) +
  geom_net(aes(colour = as.factor(drink)), size = .75, vjust = .5,
           fontsize = 6, curvature = .25, ealpha = .6, directed = T,
           arrowsize = .5, arrowgap = .01) + theme_net() + 
  scale_color_brewer(palette = "YlOrRd", name = "Drinking Frequency", labels = c("never", "about 2x/year", "about 1x/month", "about 1x/week", "More than 1x/week")) + 
  facet_wrap(~period, nrow = 1)
@

Several actor-oriented longitudinal network models can be fit to this data in RSiena. In order to determine the importance of certain actor-level variable effects included in the models, we simulated networks from different fitted models in Siena, and placed them in lineups to determine if there is any difference between the models.
}

<<fit_models>>=
# First create a 50 * 50 * 3 array composed of the 3 adjacency matrices
friendshipData <- array( c( friend.data.w1, friend.data.w2, friend.data.w3 ),
                         dim = c( 50, 50, 3 ) )
# and next give this the role of the dependent variable:
friendship <- sienaDependent(friendshipData)
alcohol <- varCovar( as.matrix(drink[,1:2]) )
# create siena data object
mydata <- sienaDataCreate( friendship, alcohol)

myeff <- getEffects( mydata )
# 3 actor effects
myeff1 <- includeEffects( myeff, egoX, altX, egoXaltX, interaction1 = "alcohol" )
# 1 actor effect 
myeff2 <- includeEffects( myeff, egoX, interaction1 = "alcohol")
# Create object with algorithm settings
myalgorithm <- sienaAlgorithmCreate( projname = 's50' )
#Estimate parameters
ans1 <- siena07( myalgorithm, data = mydata, effects = myeff1)
ans2 <- siena07( myalgorithm, data = mydata, effects = myeff2)

# create data frames with fitted parameters to use later (need to find a more automated way to do this.)
rate.params <- data.frame(rate.param = c("Rate", "Rate"),
           rate.param.val = ans1$rate)
eval.params <- data.frame(eval.param = c("density", "recip"), eval.param.val = ans1$theta[1:2])
eval.int.params <- data.frame(eval.int.param = c("altX", "egoX", "egoXaltX"), 
                              eval.int.param.val = ans1$theta[3:5])
rate.params2 <- data.frame(rate.param = c("Rate", "Rate"),
                          rate.param.val = ans2$rate)
eval.params2 <- data.frame(eval.param = c("density", "recip"), eval.param.val = ans2$theta[1:2])
eval.int.params2 <- data.frame(eval.int.param = "egoX", 
                              eval.int.param.val = ans2$theta[3])
@

<<simulation_fn, echo=FALSE>>=
SimulateNextWave <- function(init.waves, V0, 
                                     rate.params.names, rate.params.vals,
                                     eval.params.names, eval.params.vals,
                                     eval.int.params.names, eval.int.params.vals){
  if (ncol(init.waves[,,1]) != nrow(init.waves[,,1])){
    return("Error: Non-square adjacency matrix provided")
  } 
  if (length(eval.params.vals) != length(eval.params.names)){
    return("Error: Number of eval parameters to include in model does not
           match number of eval parameter values provided.")
  }
  if (length(rate.params.vals) != length(rate.params.names)){
    return("Error: Number of rate parameters to include in model does not
           match number of rate parameter values provided.")
  }
  if (length(eval.int.params.vals) != length(eval.int.params.names)){
    return("Error: Number of interaction eval parameters to include in model does not
           match number of interaction eval parameter values provided.")
  }
  init.waves <- friendshipData
  n <- ncol(init.waves[,,1])
  V0 <- drink
  c <- length(unique(V0[,1]))
  p.e <- length(eval.params.names)
  p.r <- length(rate.params.names)
  p.e.i <- length(eval.int.params.names)
  # Simulates M consecutive network and behavior waves, with n actors,
  # with c categories of the behavior variable, from initial network init.net.adj
  # according to a stochastic actor-oriented model with actor covariate V0
  # with parameter values params.values for the parameters of interest params.names
  Va <- varCovar(V0)
  X   <- sienaDependent(init.waves)
  InitData <- sienaDataCreate(X, Va)
  InitEff0 <- getEffects(InitData)
  # sink to avoid printing to the screen
  #sink("eff.txt")
  # Specify the parameters.
  # The rate parameters are first multiplied by 10,
  # which will be used only to get from the totally random network XX[,,1] = X0
  # to the network that will be the simulated first wave.

  for (i in 1:p.r){
    InitEff0 <- setEffect(InitEff0, rate.params.names[i], type = 'rate', initialValue = rate.params.vals[i], period = i, character = T)
  }
  for (i in 1:p.e){
    InitEff0 <- setEffect(InitEff0, eval.params.names[i], type = 'eval', initialValue = eval.params.vals[i], character = T)
  }
  for (i in 1:p.e.i){
    InitEff0 <- setEffect(InitEff0, eval.int.params.names[i], type = 'eval', interaction1 = "Va", initialValue = eval.int.params.vals[i], character = T)
  }

  ## The parameter given for n3 should be larger than sum(InitEff0$include)
  InitAlg <- sienaAlgorithmCreate(projname="Init", nsub=0,
                                  cond=FALSE, simOnly=TRUE)
  # Simulate the first wave.
  InitSim   <- siena07(InitAlg, data=InitData, eff=InitEff0,
                       returnDeps=TRUE, batch=TRUE, silent=TRUE)
  return(data.frame(InitSim$sims[[1000]][[1]][[1]][[1]]))
}  
@

<<simulate_network>>=
# one from 3-effect model (the "data" model)
sim_dat <- SimulateNextWave(init.waves = friendshipData, V0 = drink, 
                 rate.params.names = rate.params$rate.param, 
                 rate.params.vals = rate.params$rate.param.val,
                 eval.params.names = eval.params$eval.param,
                 eval.params.val = eval.params$eval.param.val,
                 eval.int.params.names = eval.int.params$eval.int.param,
                 eval.int.params.vals = eval.int.params$eval.int.param.val)
sim_dat_net <- merge(sim_dat, data.frame(id =1:50), by.x = "X1", by.y = "id", all = T)

# two from the 1-effect model (the "null" model)
sim_null1 <- SimulateNextWave(init.waves = friendshipData, V0 = drink,                         rate.params.names = rate.params2$rate.param,
                    rate.params.vals = rate.params2$rate.param.val,
                      eval.params.names = eval.params2$eval.param,
                     eval.params.val = eval.params2$eval.param.val,
           eval.int.params.names = eval.int.params2$eval.int.param,
        eval.int.params.vals = eval.int.params2$eval.int.param.val)
sim_null1_net <- merge(sim_null1, data.frame(id =1:50), by.x = "X1", by.y = "id", all = T)
sim_null2 <- SimulateNextWave(init.waves = friendshipData, V0 = drink, 
                       rate.params.names = rate.params2$rate.param, 
                     rate.params.vals = rate.params2$rate.param.val,
                       eval.params.names = eval.params2$eval.param,
                     eval.params.val = eval.params2$eval.param.val,             eval.int.params.names = eval.int.params2$eval.int.param,         eval.int.params.vals = eval.int.params2$eval.int.param.val)
sim_null2_net <- merge(sim_null2, data.frame(id =1:50), by.x = "X1", by.y = "id", all = T)

# null is smaller model
small_lu <- rbind(sim_dat_net, sim_null1_net,sim_null2_net)
small_lu$group <- rep(sample.int(3), c(nrow(sim_dat_net), nrow(sim_null1_net), nrow(sim_null2_net)))

ggplot(small_lu, aes(from_id = X1, to_id = X2)) + 
  geom_net(ealpha = .7, fiteach = T, directed = T, size = 1, arrowsize = .5) + 
  theme_net() + 
  facet_wrap(~group, nrow=1)
@

The ``data" model is facet \Sexpr{small_lu$group[1]}.
Notice the caption at the top of Table~\ref{vis_accept}.

\begin{table}
%% Table captions on top in journal version
 \caption{SciVis Paper Acceptance Rate: 1994-2006}
 \label{vis_accept}
 \scriptsize
 \begin{center}
   \begin{tabular}{cccc}
     Year & Submitted & Accepted & Accepted (\%)\\
   \hline
     2004 & 167 & 46 & 27.6\\
     2005 & 268 & 88 & 32.8\\
     2006 & 228 & 63 & 27.6
   \end{tabular}
 \end{center}
\end{table}

\subsection{Mezcal Head}

Citing happens only with cite - like this for the lineup protocol~\cite{buja:2009}, and for RSiena~\cite{RSiena}. Footnotes are ok%
\footnote{Footnotes appear at the bottom of the column}.

\section{Conclusion}

%% if specified like this the section will be committed in review mode
\acknowledgments{
The authors wish to thank A, B, C. This work was supported in part by
a grant from XYZ.}

\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{references}
\end{document}