### 2016-03-29 PaperThoughts.Rnw
managed to wade my way through the entire process of ML estimation for SAOMs. I think I understand it fully now.

\documentclass[a4paper]{report}
\usepackage[margin=.75in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{amsmath,amssymb,array}
\usepackage{booktabs}
\usepackage{tikz}
%% load any required packages here

\usepackage{adjustbox}            %% to align tops of minipages
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{afterpage}

\usepackage{color}
\definecolor{purple}{rgb}{.4,0,.8}
\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\st}[1]{{\color{purple} #1}}

\begin{document}

{\Large \textbf Stochastic Actor-Oriented Models for Social Networks}

\section{Terminology \& Notation}

\begin{itemize}
\item \textbf{Network} - A collection of actors and ties between them. A whole network (all of the nodes and the ties between them) is denoted $x$. In SAOMs, the networks are observed in waves for time points $m = 1, \dots, M$. Sometimes, the networks observed are denoted as $x(t_m)$.  
\item \textbf{Node} or \textbf{Actor} - The entities that form ties in a network.  The number of nodes in a network is denoted $n$. The set of nodes, $\{i : i = 1, \dots n\}$ is denoted $\mathcal{N}$. 
\item \textbf{Edge} or \textbf{Tie} - The connections between actors in a network. A tie from actor $i$ to actor $j$ is denoted by $x_{ij}$. In SAOMs, ties are almost always directed, so $x_{ij}$ does not necessarily equal $x_{ji}$. $x_{ij} = 1$ if there is a tie from actor $i$ to actor $j$ and $x_{ij} = 0 $ if there is not a tie from actor $i$ to actor $j$. Ties from an actor to itself, $x_{ii}$, are structurally zero, meaning that self-ties are not allowed. In a network of $n$ nodes, there are $n(n-1)$ possible ties. 
\item \textbf{Adjacency matrix} - The adjacency matrix of a network is a numerical representation of it. The adjaceny matrix of $x$ is denoted $\mathcal{A}(x)$, is of dimension $n \times n$ and has $ij^{th}$ entry equal to the value of $x_{ij}$ for all $i \neq j \in \mathcal{N}$. The diagonal entries of $\mathcal{A}(x)$, $ii$ for $i = 1, \dots, n$, are all structurally 0.  
\item $\mathbf{X(t)}$ - A continuous time Markov process \emph{in which} the observed panel of $M$ networks of size $n$ is embedded. For each $t = 1, 2, 3, \dots$, there is one tie change. In other words, the network $X(t+1)$ only differs from $X(t)$ by one tie, $x_{ij}$. Call this tie $x_{ij}(t)$ in network $X(t)$ and $x_{ij}(t+1)$ in network $X(t+1)$. Then, $x_{ij}(t+1) = 1-x_{ij}(t)$. Note that $t$ is different from $t_m$. Here is an illustrated example: \\

\begin{centering}
\textbf{\emph{$X(t)$, the continuous time Markov process}} \\
\begin{tikzpicture}
\draw (0,0) -- (12.5,0);
\foreach \x in {0,1,2,3,4,5,6,7,8,9,10,11,12}
\draw[shift={(\x,0)},color=black] (0pt,3pt) -- (0pt,0pt);
\foreach \x in {0,1,2,3,4,5,6,7,8,9,10,11,12}
\draw[shift={(\x,.05)},color=black] (0pt,0pt) -- (0pt,-3pt) node[above]
{$\x$};
\foreach \x in {0,1,2,3,4,5,6,7,8,9,10,11,12}
\draw[shift={(\x,.5)},color=black] (0pt,0pt) -- (0pt,-3pt) node[above]
{$X(\x)$};
\draw (0,0) node[below]{$t_1$};
\draw (4,0) node[below]{$t_2$};
\draw (7,0) node[below]{$t_3$};
\draw (10,0) node[below]{$t_4$};
\draw (12,0) node[below]{$t_5$};
\draw (0,-.5) node[below]{$x(t_1)$};
\draw (4,-.5) node[below]{$x(t_2)$};
\draw (7,-.5) node[below]{$x(t_3)$};
\draw (10,-.5) node[below]{$x(t_4)$};
\draw (12,-.5) node[below]{$x(t_5)$};
\end{tikzpicture} \\

\textbf{\emph{$x(t_m)$, the observed networks}} \\

\end{centering}

Thus, each network observation, $x(t_m)$ represents the concatenation of (usually) multiple consecutive steps in the Markov chain in SAOMs. The state space for the the Markov chain is $\mathcal{S}$, the large but finite set of all possible adjacency matrices for $n$ nodes with zeroes on the diagonal. Note that $|\mathcal{S}| = 2^{n(n-1)}$.\footnote{See http://oeis.org/A053763.} Also note that the time points $t$ are not strictly 1 unit of time apart. The space between them is dictated by the distribution of the waiting times in the Markov process. 
\item \textbf{Rate of Change} - One of the defining elements of SAOMs is the function for the rate of change of a network and of the actors in that network.  For actor $i$, the rate function is commonly denoted $\lambda_i(\alpha, x)$, which dictates how quickly actor $i$ gets opportunities to change one of its ties, $x_{ij}$. In this function, $\alpha$ is a parameter and $x$ is the current network state. The rate function for the whole network is $\lambda(\alpha, x) = \sum_i \lambda_i(\alpha, x)$. For any time point, $t$, where $X(t) \equiv x$, the waiting time to the next change opportunity by \emph{any} actor has distribution \emph{Exponential}$(\lambda(\alpha, x))$. This is because the chain of network evolutions is Markovian. There are many possibilities for the rate function. The simplest is that it is constant between observations $x(t_m)$. The rate function can also depend on covariate values of the actors or structural network elements such as outdegree. For instance, $\lambda_i(\alpha, x) = \exp\{\sum_k \alpha_k r_{ik}(x)\}$, where $r_{ik}(x)$ is a function of the current network state and the node $i$ (one of the aforementioned possible dependencies), and $k = 1, \dots K$ could be any number of functions as chosen by the researcher from previous subject matter knowledge or otherwise.  
\item \textbf{Probability of Change (nodes)} - Given that a change, call it $X(t) \equiv x \rightarrow X(t+1)$ occurs, each node has its own probability of changing. Recall that from time $t$ to time $t+1$, only one tie is changing (call it $x_{ij}$), so that only one node, $i$, is given the power to change one of its $n-1$ available ties from $x_ij$ to $1-x_{ij}$. The probability that node $i$ will be the one to change a tie, given $x \rightarrow X(t+1)$ is $\pi_i(\alpha, x) \equiv \pi_i(\alpha, x|x\rightarrow X(t+1)) = \frac{\lambda_i(\alpha, x)}{\lambda(\alpha, x)}$.
\item \textbf{Permitted Changes} - First note a slight abuse of notation. The current state of the network, which was previously denoted $X(t)$ and $x$, will now be denoted $x^0$. Then, let $x$ instead represent the new, as yet undetermined network $X(t+1)$. So, given that a change, $x^0 \rightarrow x$ will occur, and that node $i$ has been selected to change a tie, there is a set of permitted changes, $\mathcal{A}_i(x^0) = \{\mathcal{A}(x^0) \cup \mathcal{A}_i^r(x^0)\}$, where $\mathcal{A}(x^0)$ is the current network's adjacency matrix, and $ \mathcal{A}_i^r(x^0) = \{x | x_{ij} = 1 - x_{ij}^0 \text{for only one } j \in \mathcal{N},  j \neq i, \}$.  So, the set of permitted changes includes the set of all adjacency matrices with all entries equivalent to the current adjacency matrix with the exception of one of the non-fixed, $n-1$ entries of the $i^{th}$ row and the current adjacency matrix, meaning that $|\mathcal{A}_i(x^0)| = n$. Additionally, define the complete set of possible changes from $x^0 \rightarrow x$ as $\mathcal{A}_{1:n}(x^0) = \{\mathcal{A}(x^0) \cup \mathcal{A}_1(x^0) \cup \mathcal{A}_2(x^0) \cup \cdots \cup \mathcal{A}_n(x^0)\}$. This is the set of all networks that differ from $x^0$ by only one tie and has size $|\mathcal{A}_{1:n}(x^0)| = n(n-1) + 1$. 
\item \textbf{Objective function} - So, once a change in node $i$ is given, \emph{how} does that node determine which of the $n$ permitted changes it will make? Its choice is determined in large part by its objective function, $f_i(\beta, x^0, x)$, where $\beta$ is a parameter. This function can be interpreted as the ``relative attractiveness" for node $i$ associated with making a change from the current network state, $x^0$ to a new network state, $x$, where $x \in \mathcal{A}_i(x^0)$. It is assumed that actors want to maximize this function when given an opportunity for change. For example, if changing from $x^0$ to $x$ will decrease the value of $f_i$ for all $x \in \mathcal{A}_i(x^0)$, the node will be very likely to \emph{not} change any ties. Many models don't include the current state, $x^0$, in the objective function, but instead only consider $x$, the future state, and just attempt to get closer to their optimum state regardless of where they currently are. The objective function is usually defined linearly as $f_i(\beta, x^0, x) = \sum_k \beta_k s_{ik}(x^0, x)$, where the functions $s_{ik}$ are determined from subject matter knowledge. These, like the $r_{ik}$ in the rate function, can be actor covariates or structural variables like outdegree or reciprocated ties. 
\item \textbf{Probability of Change (tie given node)} - Given the change of a tie from node $i$, the probability of the change $x^0 \rightarrow x$, where $x \in \mathcal{A}_i(x^0)$ is $p_i(\beta, x^0, x) \equiv p_i(\beta, x^0, x | x^0\rightarrow x, x \in \mathcal{A}_i(x^0)) = \frac{\exp \{f_i(\beta, x^0, x)\}}{\sum_{\tilde{x}} \exp \{f_i(\beta, x^0, \tilde{x})}$, where $\tilde{x} \in \mathcal{A}_i(x^0)$. For a specific tie, $x_{ij}$, with $j \neq i$, this expression can be re-written as $p_{ij}(\beta, x^0) = \frac{\exp \{f_i(\beta, x^0, x^j)\}}{f_i(\beta, x^0, x^0) + \sum_{j \neq i} \exp \{f_i(\beta, x^0, x^j)} $, where $x^j$ has ties $x^j_{ih} = x^0_{ih}$ for all $h \neq j$ and $x^j_{ih} = 1- x^0_{ih}$ for $h = j$. Using this new expression, we can write the probability of no change for node $i$ ($p_i(\beta, x^0, x^0)$ in the original notation) as $p_{i0} \equiv p_i(\beta, x^0, x^0) = 1 - \sum_{j \neq i} p_{ij}(\beta, x^0)$. 
\item \textbf{Intensity Matrix} - The intensity matrix, $Q$, of a continuous time Markov process describes the rate of change between states of the chain. For networks, there are a very large number of possible states: $2^{n(n-1)}$. But, due to the properties of continuous-time Markov processes, there are only $n$ possible states given the current state, $n-1$ of which are uniquely determined by the node $i$ that is given the opportunity to change. Thus, the intensity matrix $Q$ is a very sparse $2^{n(n-1)} \times 2^{n(n-1)}$ matrix, with only $n(n-1) + 1$ non-zero entries in each row. Note that $n(n-1)$ of these represent the possible states that are one edge different from a given state, and the additional non-zero entry is for the state to remain the same. All other entries in a row are zero because those column states cannot be reached from the row state by just one change. The entries of $Q$ are defined as follows: let $y \neq z \in \{1, 2, \dots, 2^{n(n-1)} \}$ be indices of two different possible states of the network, $x^y, x^z \in \mathcal{S}$. For a timepoint $t$ define $x^y = X(t)$ and let $x^z \in \mathcal{A}_i(x^y)$ for each $i \in \mathcal{N}$. Then the $yz^{th}$ entry of $Q$ is:
\[ q(x^y, x^z) = \begin{cases} 
      \lambda_i(\alpha, x^y)p_i(\beta, x^y, x^z) & \text{if } x^z \in \mathcal{A}_i(x^y), \forall i \in \mathcal{N} \\
      0 & \text{if } x^z \notin \mathcal{A}_i(x^y) \text{ for any i } \in \mathcal{N}
   \end{cases}
\]
Properties of intensity matrices dictate that the diagonal entry, where the state doesn't change, is defined as $q(x^y,x^y) = -\sum_{z\neq y} q(x^y, x^z)$ so that the row sum is 0. Thus, the rate of change between any two states that differ by only one tie from $i \to j$ is the product of the rate at which actor $i$ gets to change a tie and the probability that the tie that will change is the tie to node $j$.\footnote{Just to be clear, the change is from $x^y_{ij}$ to $x^z_{ij} = 1 - x^y_{ij}$.} Furthermore, the theory of continuous time Markov chains gives that the matrix of transition probabilities between observation times $t_{m-1}$ and $t_{m}$ is dependent only on the difference between timepoints, $t_m - t_{m-1}$ and is equal to $e^{(t_m - t_{m-1})Q}$, where $Q$ is the matrix defined above and $e^X$ for a real or complex square matrix $X$ is equal to $\sum_{k=0}^{\infty} \frac{1}{k!} X^k$. There is a package in \texttt{R} to do this matrix exponential operation, but I can't imagine doing this for networks for any $n > 5$.
\item \textbf{Moving from $x(t_{m-1})$ to $x(t_m)$} - First define $x(t_1)$ as the first network observed and $x(t_M)$ as the last network observed. Networks in between are written $x(t_{m})$ for $m = 2, \dots, M-1$. The observations $x(t_1), x(t_2), \dots, x(t_M)$ are embedded in the continuous-time Markov chain $\{X(t)|t_1 \leq t \leq t_M\}$. Moving from the observation $x(t_{m-1})$ to the observation $x(t_m)$ requires $R$ changes or steps in the Markov process $X(t)$ where $R = \sum_{i,j \in \mathcal{N}} |x_{ij}(t_m) - x_{ij}(t_{m-1})|$. Note that the observed network change from $x(t_{m-1})$ to $x_(t_m)$ is conditionally independent of all other network observations given $x(t_{m-1})$. Denote the $R$ timepoints in the the Markov process that correspond to the unobserved changes in the network as $T_1, T_2, \dots, T_{R-1}$ and define $T_0 = t_{m-1}$ and $T_R \leq t_{m} < T_{R+1}$. For each timepoint $T_r$, for $r = 1, \dots R$, there is a single actor, denoted $I_r$ that gets an opportunity to change at this timepoint. This actor changes one of its ties to one node, $J_r$. If there is no change, $I_r = J_r$. Formally, the pair $(I_r,J_r)$ is the only $(i,j)$ pair for which $x_{ij}(T_{r-1}) \neq x_{ij}(T_r)$ if such a pair exists, and $(I_r, J_r) = (I_r, I_r)$ otherwise. The triplet $(T_r, I_r, J_r)$ forms a stochastic process for $r = 1, \dots, R$ that, given the prior observation, $x(t_{m-1})$ for $m = 2, \dots, M$, completely determines $x(t)$ for $t_{m-1} < t < t_m$. 
\item \textbf{Augmenting Data} (see section on ML estimation) - The augmenting data consist of $R$ and the pairs $(I_r, J_r)$ for $r= 1, \dots, R$. (Note the timepoints $T_r$ are left out of the augmenting data.) The possible outcomes of the augmenting data are determined by the condition that if for a given tie, $x_{ij}(t_{m-1})$, the number of timepoints $1 \leq r\leq R$ at which $(i_r, j_r) = (i,j)$ is even if $x_{ij}(t_{m}) = x_{ij}(t_{m-1})$ and is odd if $x_{ij}(t_{m}) \neq x_{ij}(t_{m-1})$. 
\item \textbf{Sample Path} - The sample path is the stochastic process $V = ((I_r, J_r):r = 1, \dots, R)$. Note that the elements of $V$ where $I_r = J_r$ are redundant. However, they are kept in the process because they help with computation of the likelihood. For notational convenience, define $x^{(r)} = x(t_r)$. Using this notation, write $x(t_{m-1}) = x^(0)$ for the independent cases where $m = 2, \dots, M$. Then, the networks $x^{(r)}$ and $x^{(r-1)}$ only differ in the element $(I_r, J_r) \in V$, provided that $I_r\neq J_r$.
\item \textbf{PDF of Sample Path} - The sample path has probability distribution function 
  $$ p_{sp}(V|\alpha, \beta, x^{(0)} ) = Pr(T_R \leq t_m < T_{R+1} | x^{(0)}, V, \alpha, \beta) \times \prod_{r=1}^R \pi_{i_r}(\alpha, x^{(r-1)})p_{i_r,j_r}(\beta,x^{(r-1)}). $$
The first component, $Pr(T_R \leq t_m < T_{R+1} | x^{(0)}, V, \alpha, \beta)$ is the probability that your next network observation at time point $t_m$ comes \emph{before} the next change in the continuous time Markov chain, $X(T_{R+1})$. From a discussion that follows, $Pr(T_R \leq t_m < T_{R+1} | x^{(0)}, V, \alpha, \beta)$, the probability that your next network observation at time point $t_m$ comes \emph{before} the next change in the continuous time Markov chain, $X(T_{R+1})$, can be approximated by 
$$ \kappa(\alpha, x(t_{m-1}), V) \approx \int_{t_{m-1}}^{t_m} p_{T_R}(s) \text{Pr}(\{T_{R+1} - T_R > t_{m} - T_R \} | T_R = s) ds $$
$T_{R+1}$ has distribution Exponential$(\lambda(\alpha, x^{(R)}))$. So, Pr$(\{T_{R+1} - T_R > t_m - T_R \}) = \text{Pr}(T_{R+1} > t_m) = 1 - \text{Pr}(T_{R+1} < t_m) = 1 - (1 - \exp\{-\lambda(\alpha, x^{(R)})t_m\}) = \exp\{-\lambda(\alpha, x^{(R)})t_m\}$. So, the above integral can be written as   
$$ \kappa(\alpha, x(t_{m-1}), V) \approx \int_{t_{m-1}}^{t_m} p_{T_R}(s) \exp\{-\lambda(\alpha, x^{(R)})(t_m-s)\}. ds $$
This quantity is approximately equal to $\frac{p_{T_R}(t_m)}{\lambda(\alpha, x^{(R)})}$. The second expression, $\prod_{r=1}^R \pi_{i_r}(\alpha, x^{(r-1)})p_{i_r,j_r}(\beta,x^{(r-1)})$, is the PDF of the stochastic process $V$. So, the probability of the sample path can be approximated by:
$$ p_{sp}(V|\alpha, \beta, x^{(0)}) \approx \frac{p_{T_R}(t_m)}{\lambda(\alpha, x^{(R)})} \times \prod_{r=1}^R \pi_{i_r}(\alpha, x^{(r-1)})p_{i_r,j_r}(\beta,x^{(r-1)}). $$
So, for observed data, augmented by the sample path, ``the likelihood conditional on $x(t_{m-1})$ can be expressed directly, either exactly [(for constant $\lambda_i$)] or in good approximation."
\item \textbf{Distribution of waiting times between $x(t_{m-1})$ and $x(t_m)$} - Let $T_{r+1} - T_r$ be the waiting time between one of the $r = 1, \dots, R$ changes between $x(t_{m-1})$ and $x(t_m)$. After conditioning on $x(t_{m-1}) \equiv x^{(0)}$ and $V$, these waiting times have distribution $T_{r+1} - T_{r} \sim \text{Exponential}(\lambda(\alpha, x^{(r)}))$ for $r = 0, 1, \dots, R-1$. Following from this, the distribution of $T_R - t_{m-1}$, the time it takes for the network to move from state $x(t_{m-1})$ to state $x(t_{m})$ is the convolutions of the Exponential$(\lambda(\alpha, x^{(r)}))$ distributions for $r = 0, \dots, R-1$. The pdf of $T_R - t_{m-1}$ is\footnote{Source: \url{http://www.ccms.or.kr/data/pdfpaper/jcms21_4/21_4_501.pdf}}: 
$$ h(T_R - t_1 | \alpha, x(t_{m-1}), V) = \sum_{r=0}^{R-1} \frac{\lambda(\alpha, x^{(0)})\cdot \lambda(\alpha, x^{(1)}) \cdot \lambda(\alpha, x^{(R-1)})}{\prod_{q\neq r, q = 0}^{R-1} (\lambda(\alpha, x^{(q)}) - \lambda(\alpha, x^{(r)}))} \cdot \exp\{-(T_R - t_{m-1})\cdot \lambda(\alpha, x^{(r)})\}$$
In the special case that the $\lambda_{i}(\alpha, x)$ rates are constant, call it $\alpha_1$, then the number of steps in the Markov chain to move from $x(t_{m-1})$ to $x(t_m)$, $R$ has distribution $R \sim \text{Poisson}(n\alpha_1(t_{m} - t_{m-1}))$. In this case, $Pr(T_R \leq t_m < T_{R+1} | x^{(0)}, V, \alpha, \beta) = \kappa(\alpha_1, x^{(0)}, V) = \exp\{-n\alpha_1(t_{m} - t_{m-1})\}\frac{(n\alpha_1(t_m - t_{m-1}))^R}{R!}$. However, when the individual rates are different for each actor, this probability must be approximated. 
\item \textbf{Distribution of $T_R$} - $T_R$ is the timepoint in the Markov chain at which the state $x(t_m)$ is acheived. Since $T_R - t_{m-1}$ is a convolution of exponentials with rate parameters $\lambda(\alpha, x^{(r)})$ for $r = 0, R-1$, by the Lyapounov Central Limit Theorem, the distribution of $T_R$ is approximately normal with mean $\mu_{\alpha}$ and variance $\sigma^2_{\alpha}$ where 
$$ \mu_{\alpha} = \sum_{r=0}^{R-1} \frac{1}{\lambda(\alpha, x^{(r)})} \quad \text{and} \quad  \sigma^2_{\alpha} = \sum_{r=0}^{R-1} \frac{1}{\lambda(\alpha, x^{(r)})^2} $$
The PDF of $T_R$ is thus $p_{T_R} \approx \frac{1}{\sqrt{2\pi \sigma^2_{\alpha}}} \exp\left\{\frac{-((t-t_{m-1})-\mu_{\alpha})^2}{2\sigma^2_{\alpha}}\right\}$. 
\end{itemize}

\section{Important Facts about SAOMs}
\begin{itemize}
\item In SAOMs, the observed networks in a longitudinal study, call them $x_{t_1}, \dots, x_{t_M}$, are discrete observations embedded in a continuous time process, $X(t)$. This means that the probability model is defined \emph{independently} of the observational design. Thus, the time between observations can be irregular and still be analyzed under this model.
\item The continuous-time Markov process $\{X(t) | t_1 \leq t \leq t_M\}$ has a distribution that consitutes an exponential family, which are of the form $f(y|\theta) = \exp\{\sum_{j=1}^s q_j(\theta) T_j(y) \} c(\theta) h(y)$ where $y$ is the data vector and $\theta$ is the scalar or vector parameter. This means the SAOMs can be viewed as ``incompletely observed exponential family" models. 
\item Inference is computer-intensive and time-consuming because of ``elaborate MCMC procedures" that must be implemented.
\item The definition of SAOMs means that data can be simulated directly from the probability distribution given the initial network state $x_{t_1}$. 
\item However, SAOMs are ``too complicated for the calculation of likelihoods or estimators in closed form." The distribution, however, can be easily simulated from, so that estimation can proceed by MCMC simulation.
\item ML estimation of SAOMs is based on the missing data principle. The ``missing data" here are the paths from $x(t_{m-1})$ to $x(t_{m})$ for all $m = 2, \dots, M$. This set can be denoted for each $m$ by $V_m = \{(I_{m1}, J_{m1}),\dots, (I_{mR_m}, J_{mR_m})\}$. So, the complete set of augmenting data is $V = (V_2, \dots V_M)$.  
\end{itemize}

\section{Maximum Likelihood Estimation}
Sketch of Algorithm:
\begin{enumerate}
\item Use the first network observation, $x(t_1)$, as a starting value. All analysis proceeds conditioning on $x(t_1)$.  
\item For $m = 2, 3, \dots, M$ where $M$ is the total number of observed networks, augment the observed data with random draws from the sequence of intermediate steps in the Markov process, $x(t)$, that could have led from $x(t_{m-1})$ to $x(t_m)$. i.e. draw from the set {\small $$\{(x(t^*_1), x(t^*_2), \dots x(t^*_{C-1}))' | x(t^*_1) \in \mathcal{A}_{1:n}(x(t_{m-1})) \& x(t^*_2) \in \mathcal{A}_{1:n}(x(t^*_1)) \& x(t^*_3) \in \mathcal{A}_{1:n}(x(t^*_2)) \& \cdots \& x(t_{m}) \in \mathcal{A}_{1:n}(x(t^*_{C-1})) \}$$ } where $C = \sum_{i,j \in \mathcal{N}} |x_{ij}(t_m) - x_{ij}(t_{m-1})|$. Note that $C$ in the number of single edge changes needed to get from $x(t_{m-1})$ to $x(t_m)$, so $C-1$ is the number of intermediate networks. These draws can be simulated using a Metropolis-Hastings algorithm. 
\item Use the draws from 2 as updates in a Robbins-Monro algorithm (?) to find the solution of the likelihood equation. 
\end{enumerate}

Full Iterative Algorithm: For each iteration $N = 1, 2, \dots$
\begin{enumerate}
\item For each $m = 2, \dots, M$, make a large number of Metropolis-Hastings steps of the following form.  Let $\underline{v} = ((i_1, j_1), \dots, (i_R, i_R))$ be a given path from $x(t_{m-1})$ to $x(t_m)$ in the whole possible set of paths $\mathcal{V}$. Then propose a new path, $\tilde{\underline{v}}$ from the proposal distribution that consists of all of the possible small changes: 
  \begin{enumerate}
  \item ``Paired Deletions" - Of all pairs of indices $r_1, r_2$ such that $(i_{r_1}, j_{r_1}) = (i_{r_2}, j_{r_2})$, $i_{r_1} \neq j_{r_1}$, randomly select a pair $(r_1, r_2)$ and delete both from the current path.
  \item ``Paired Insertions" - Randomly select an edge $(i,j) \in \mathcal{N}^2$ with $i \neq j$. Randomly choose 2 indices, $r_1, r_2$. Insert $(i,j)$ immediately before $r_1$ and $r_2$.  
  \item ``Single Insertions" - At a random place in the current path, insert $(i,i)$ for a random $i \in \mathcal{N}$.
  \item ``Single Deletions" - Of all elements in the current path that satisfy $i_r = j_r$, randomly delete one of them. 
  \item ``Permutations" - For randomly chosen $r_1 < r_2$, where $r_2 - r_1$ is bounded from above by some smallish number to avoid lengthy computation (tuning parameter?), the segment of consecutive elements $(i_r,j_r)$, $r = r_1, \dots, r_2$ is randomly permuted. 
  \end{enumerate}
Denote the proposal probabilities $u(\tilde{\underline{v}} | \underline{v})$ and the target probabilities $p(\underline{v})$. The target distribution is $p(\underline{v}) \propto p_{SP}(\underline{v} | \alpha, \beta, x(t_{m-1}))$. Then the acceptance probability for a proposal path $\tilde{\underline{v}}$ and a current path $\underline{v}$ is: $\min \left\{ 1, \frac{p(\tilde{\underline{v}})u(\underline{v}|\tilde{\underline{v}})}{p(\underline{v})u(\tilde{\underline{v}}| \underline{v})}\right\}$. The series of M-H steps will result in a new set of augmenting data, $(v^{(N)}) = (v_2^{(N)}, \dots, v_M^{(N)})$. 
\item Compute 
  $$ S_{XV}(\hat{\theta^{(N)}}; x, v^{(N)}) = \sum_{m=2}^M S_m(\hat{\theta^{(N)}}; x(t_{m-1}), v_m^{(N)}) $$
$S_{XV}(\hat{\theta^{(N)}}; x, v^{(N)})$ is the complete data score function. $S_m(\hat{\theta^{(N)}}; x(t_{m-1}), v_m^{(N)}) = \frac{\partial \log p_m(v_m; \theta | x(t_{m-1}))}{d\theta}$ is the total data score function at step $m$. $p_m$ is $p_{SP}(v_m; \theta = (\alpha, \beta)' | x(t_{m-1}))$, the probability of the sample path $v_m$. 
\item Update $$\theta^{(N+1)} =(\alpha^{(N+1)}, \beta^{(N+1)}) = \theta^{(N)} + a_N D^{-1} S_{XV}(\hat{\theta^{(N)}}; x, v^{(N)})$$
where $D$ is a MC estimate of the complete data observed Fisher information matrix, $D_{XV}(\theta) = -\frac{\partial S_{XV}(\theta: x, V)}{\partial \theta}$ estimated for $\theta = \theta^{(1)}$. (i.e. evaluated at the starting value of $\theta$.) The starting value is the method of moments estimator, $\hat{\theta}^{(1)}$. (See Snijders 2001 for how to get the MoM estimate). The sequence $a_N$ is just a sequence that approaches 0 as $N \to \infty$. (I'm not sure exactly what? Maybe just $\frac{1}{N}$ would work? There's no specifics in Snijders 2001, and I'm looking through the RSiena manual for clues...). 
\end{enumerate}

\end{document}