### 2016-04-20 lineups.Rnw
add more model description

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts, mathrsfs,fancyhdr,syntonly,lastpage,hyperref,enumitem,graphicx, lscape, forloop, url}

\hypersetup{colorlinks=true,urlcolor=black}

\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\st}[1]{{\color{red} #1}}

\begin{document}
\lhead{Graph Lineups}
\chead{}
\rhead{Samantha Tyner}

\tableofcontents

\section{Introduction}

Stochastic actor-oriented models (SAOMs) are a set of network models that consider both network and actor effects when determining how and why a network changes over time. Network effects, like reciprocity and outdegree, only consider the ties between actors as important to network change, while actor effects allow for covariate values associated with each actor to influence ties between actors. In the SAOM literature, a network is denoted by $x(t_m)$, has $n$ actors, and for two actors $i \neq j \in \{1, 2, \dots, n\}$, the relationship between them is denoted by $x_ij$. This relationship $x_{ij}$ is equal to 1 if there is a tie between actors $i$ and $j$ and it is 0 if there is not a tie between them. The time component, $t_m$, represents the $m = 1, 2, \dots, M$ points of observation of the network. These networks $x(t_1), \dots, x(t_M)$ are modeled as observations of a continuous-time Markov chain, where most of the steps in the chain are unobserved.

\par Between any two discrete time points, $t_{m}$ and $t_{m-1}$ for $m = 2, \dots, M$, each actor gets the opportunity to change one of their ties at a rate of $\alpha_{m-1}$. This is due to the properties of the Markov chain, and it also assumes that the rate of change is constant for each actor between $t_{m-1}$ and $t_{m}$. When an actor is given an opportunity to change, it tries to maximize its objective function, $f_i(\beta, x)$, where $x$ is the potential new network state that actor $i$ can reach by changing one of its ties, $x_{ij}$. In model estimation, the parameters to be estimated are $\alpha_1, \dots, \alpha_{M-1}$ and the (possibly vector-valued) $\beta$

\section{Models}

We fit \st{three} different stochastic actor-oriented models to a subset of the 50 actor dataset from the ``Teenage Friends and Lifestyle Study" that is provided on the RSiena webpage.\footnote{\url{http://www.stats.ox.ac.uk/~snijders/siena/s50_data.htm}} We chose to subset the data to decrease the cognitive load on our experiment's subjects. The subset contained actors 20 through 35 and the ties between them, as well as the drinking behavior of each actor at each of the three waves. This specific subset was chosen because it showed somewhat higher connectivity than other subsets, as we've emphasized in the visualizations of the three network adjacency matrices below. For model fitting, we condition on wave 1 and estimate the parameters of our models from the second and third waves.

<<get_sm_friends, echo=FALSE, fig.width=8, fig.height=3, message = FALSE, warning = FALSE>>=
source("Code/small_friends.R")
ggplot(data= view_rs2, aes(x = x, y=y)) + 
  geom_tile(aes(fill = as.factor(value))) + scale_fill_manual(values = c('white', 'black')) +
  geom_rect(data= NULL, inherit.aes = FALSE, 
            aes(xmin = 20, xmax = 35, ymin = 20, ymax = 35, fill =NA), color = 'red') + 
  facet_wrap(~wave) + theme_bw() + 
  theme(aspect.ratio=1)

@

The first wave of the network, which is conditioned on in estimation, is given below.

<<wave1, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3>>=
library(sna)
library(network)
actual1 <- merge(data.frame(as.edgelist(as.network(fd2.w1))), 
                 data.frame(id = 1:16, drink = drink2[,1]), 
                 by.x = "X1", by.y = "id", all = T)
for (j in 1:nrow(actual1)){
      if (!(actual1$X1[j] %in% actual1$X2) & is.na(actual1$X2[j])){
        actual1$X2[j] <- actual1$X1[j]
      } else {actual1$X2[j] <- actual1$X2[j]}
}
actual2 <- merge(data.frame(as.edgelist(as.network(fd2.w2))), 
                 data.frame(id = 1:16, drink = drink2[,2]), 
                 by.x = "X1", by.y = "id", all = T)
for (j in 1:nrow(actual2)){
      if (!(actual2$X1[j] %in% actual2$X2) & is.na(actual2$X2[j])){
        actual2$X2[j] <- actual2$X1[j]
      } else {actual2$X2[j] <- actual2$X2[j]}
    }
actual1$wave <- 1
actual2$wave <- 2

waves <- rbind(actual1, actual2)
library(geomnet)
ggplot(data = actual1, aes(from_id = X1, to_id = X2)) + 
  geom_net(label = TRUE, hjust = -.5, labelcolour = 'black', fiteach = T,
           aes(color = as.factor(drink))) + 
  theme_net() 
#ggplot(data = waves, aes(from_id = X1, to_id = X2)) + 
#  geom_net(label = TRUE, hjust = -.5, labelcolour = 'black', fiteach = T,
#           aes(color = as.factor(drink))) + 
#  theme_net() + theme(panel.background = element_rect(fill = "white", color = 'black')) + facet_wrap(~wave)
@

<<modelfit, echo=FALSE, results = 'hide', message=FALSE, warning = FALSE, cache = T>>=
source("Code/complete_lineup_creation.R")
lineup1 <- create_smfriend_lu(null_eff_struct = null_model_eff2, test_eff_struct = eff_models_smallFriends[[39]], M = 3)
@

The two models we fit are a ``null model" (model $M_1$) and two ``alternative models" (model $M_2$ and $M_3$). The null model only includes the default parameters that are included in the RSiena estimation. The alternative model $M_2$ includes one additional covariate parameter that was determined to be significant by the Wald test in the RSienaTest package, while the alternative model $M_3$ includes one additional structural parameter whose significance was determined in the same way. Details on these effects are given in Table~\ref{tab:models}.

<<readsimu, echo=FALSE>>=
 nulls <- read.csv("Data/distribution_null_model.csv")
# names(nulls)[-1] <- c("alpha1", "alpha2", "beta1", "beta2")
# nulls$Sim <- 1:nrow(nulls)
# nulls$Model <- "M1"
# 
 alt <- read.csv("Data/distribution_jumpTT_model.csv")
# names(alt)[-1] <- c("alpha1", "alpha2", "beta1", "beta2", "beta3")
# alt$Sim <- 1:nrow(alt)
# alt$Model <- "M2"
# 
 alt2nd <- read.csv("Data/distribution_dblpairs_model.csv")
# names(alt2nd)[-1] <- c("alpha1", "alpha2", "beta1", "beta2", "beta4")
# alt2nd$Sim <- 1:nrow(alt2nd)
# alt2nd$Model <- "M3"
# 
# alt2 <- gather(alt, parameter, estimate, 2:6)
# null2 <- gather(nulls, parameter, estimate, 2:5)
# alt2nd2 <- gather(alt2nd, parameter, estimate, 2:6)
# 
# simu <- rbind(alt2[,-1], null2[,-1])
# write.csv(simu, "Data/simulation-1000-M1-M2.csv", row.names=FALSE)

simu2 <- read.csv("Data/simulation-1000-M1-M2-M3.csv")
means <- simu2 %>% group_by(Model, parameter) %>% summarize(
  mean = mean(estimate)
)
library(dplyr)
null_mod_sv <- as.numeric(nulls[,-c(1,6,7)] %>% summarise_each(funs(mean)))
alt_mod_sv <- as.numeric(alt[,-c(1, 7,8)] %>% summarise_each(funs(mean)))
alt_mod2_sv <- as.numeric(alt2nd[,-c(1,7,8)] %>% summarise_each(funs(mean)))
@

\begin{table}[h]
\caption{\label{tab:models}Parameters and estimates of models $M_1$, $M_2$, and $M_3$. Estimates are the mean of 1000 iterations of the model estimates. The lineups that follow are simulated from models using these values.}
\centering
\scalebox{0.8}{
\begin{tabular}{lccrrr}
Effect name & Parameter & Corresponding Statistic & $M_1$ Estimate & $M_2$ Estimate & $M_3$ Estimate\\
\hline
\hline
Rate 1 (wave 1 $\rightarrow$ 2) & $\alpha_1$ & $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_2) - x_{ij}(t_1))^2 $ & 
\Sexpr{round(null_mod_sv[1], 2)} &
\Sexpr{round(alt_mod_sv[1], 2)} & 
\Sexpr{round(alt_mod2_sv[1], 2)} 
\\
Rate 2 (wave 2 $\rightarrow$ 3) & $\alpha_2$ & $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_3) - x_{ij}(t_2))^2 $ & 
\Sexpr{round(null_mod_sv[2], 2)} &
\Sexpr{round(alt_mod_sv[2], 2)} & 
\Sexpr{round(alt_mod2_sv[2], 2)}
\\
Outdegree & $\beta_1$ & $s_{i1}(x) = \sum\limits_{j=1}^n x_{ij}$ & 
\Sexpr{round(null_mod_sv[3], 2)} &
\Sexpr{round(alt_mod_sv[3], 2)} & 
\Sexpr{round(alt_mod2_sv[3], 2)}
\\
Reciprocity & $\beta_2$ & $s_{i2}(x) = \sum\limits_{j=1}^n x_{ij}x_{ji}$ & \Sexpr{round(null_mod_sv[4], 2)} &
\Sexpr{round(alt_mod_sv[4], 2)} & 
\Sexpr{round(alt_mod2_sv[4], 2)}
\\
Jumping Transitive Triplets & $\beta_3$ & $s_{i3}(x) = \sum\limits_{\forall j\neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$ & -- & 
\Sexpr{round(alt_mod_sv[5], 2)} & -- \\
\# doubly achieved distances & $\beta_4$ & $s_{i4}(x) = |\{j : x_{ij} = 0, \sum\limits_h x_{ih}x_{hj} \geq 2\}|$ & -- & -- &
\Sexpr{round(alt_mod2_sv[5], 2)}
\end{tabular}}
\end{table}

\begin{figure}
\centering
<<hist-estimates, dependson='read-simu', echo=FALSE, fig.width=8, fig.height=5, out.width='.8\\linewidth'>>=
qplot(data=simu2, estimate, fill=Model, geom="density", alpha=I(0.65)) + facet_wrap(~parameter, scales="free") + theme_bw()
@
\caption{\label{fig:hist-simulations}Histogram of the distribution of model parameters based on 1,000 simulation runs. Model parameter $\beta_3$ for jumping transitive triplets in model $M_2$ is significantly different from zero, but its inclusion also leads to significant changes in the  other model parameters of model $M_1$. The parameter $\beta_4$ for doubly acheived distances is also significantly different from zero, but has larger variance. The inclusion of $\beta_4$ also changes the estimates of the other model parameters, but not as much as the inclusion of $\beta_3$.}
\end{figure}

After estimating the parameters in each of these models, we simulate from them to obtain realizations of each of the models. The objective function for each actor in each model is given below. 
  \begin{align*}
  f^{M_1}_{i}(x) & = \hat{\beta}_1^{M_1}s_{i1}(x) +  \hat{\beta}_2^{M_1}s_{i2}(x) \\
  f^{M_2}_{i}(x) & = \hat{\beta}_1^{M_2}s_{i1}(x) +  \hat{\beta}_2^{M_2}s_{i2}(x) + \hat{\beta}_3^{M_2}s_{i3}(x) \\
  f^{M_3}_{i}(x) & = \hat{\beta}_1^{M_3}s_{i1}(x) +  \hat{\beta}_2^{M_3}s_{i2}(x) + \hat{\beta}_4^{M_3}s_{i4}(x) 
  \end{align*}
The rate parameters, $\alpha_1$ and $\alpha_2$ represent how many opportunities for change each actor gets when moving from wave 1 to 2 and from wave 2 to 3, respectively. The outdegree parameter, $\beta_1$, represents how likely an actor is to change outgoing ties. If the estimate, $\hat{\beta}_1$, is positive, the actor is more likely to create outgoing ties, while a negative estimate leads the actor to deleting outgoing ties. This effect is highly correlated with the reciprocity parameter, $\beta_2$. A negative estimate of this parameter implies that the actor is discouraged from reciprocating its incoming ties, while a positive estimate implies that the actor is encouraged to reciprocate all ties. The additional parameter in $M_2$, $\beta_3$, is a covariate parameter. The covariate in this model is the drinking behaviour of the 16 students in our data. The possible values are 1, 2, 3, and 4, which means the student drinks never, once or twice a year, once a month, or once a week. This jumping transitive triplet effect impacts the transitive closure of actors from different groups. Thus, a positive estimate encourages transitive closure when one of three actors is in a different covariate group than the other two, while a negative estimate discourages this behavior. An example of this type of closure is given in Figure~\ref{fig:jtt}. Finally, the doubly acheived distances effect is defined by the number of actors to whom actor $i$ is not directly tied, and tied through twopaths via at least two intermediaries. This is a structural effect, like the density and repciprocity effects. A positive value encourages indirect ties, while a negative value discourages the formation of indirect ties.

<<jtt, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center', fig.cap="A realization of a jumping transitive triplet, where $i$ is the focal actor, $j$ is the target actor, and $h$ is the intermediary. The group of the actors is represented by the shape of the node.">>=
jTTe <- data.frame(from = c('i', 'i', 'h'), to = c('h', 'j', 'j'))
jTTn <- data.frame(id = letters[8:10], group = c(1,1,2))

jTT <- merge(jTTe, jTTn, by.x = 'from', by.y = "id", all = T)

set.seed(12345)
ggplot(data = jTT, aes(from_id = from, to_id = to)) + 
  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='blue',vjust = -1, hjust =2, arrowgap = .05, colour = 'black') + 
  theme_net() + theme(legend.position = "none")
@

\section{Model $M_1$ versus $M_2$ lineup size 3}
\newcounter{k}
\forloop{k}{1}{\value{k} < 21}{smallfriends-m-3-rep-\arabic{k}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-3-rep-\arabic{k}} \\ }

\section{Model $M_1$ versus $M_2$ lineup size 6}
\newcounter{k1}
\forloop{k1}{1}{\value{k1} < 21}{
Lineup-Images/pdfs/smallfriends-m-6-rep-\arabic{k1}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-6-rep-\arabic{k1}} \\ }

\section{Model $M_1$ versus $M_2$ lineup size 9}
\newcounter{k2}
\forloop{k2}{1}{\value{k2} < 21}{Lineup-Images/pdfs/smallfriends-m-9-rep-\arabic{k2}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-9-rep-\arabic{k2}} \\ }

\section{Model $M_1$ versus $M_2$ lineup size 12}
\newcounter{k3}
\forloop{k3}{1}{\value{k3} < 21}{Lineup-Images/pdfs/smallfriends-m-12-rep-\arabic{k3}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-12-rep-\arabic{k3}} \\ }

\section{Model $M_1$ versus $M_2$ lineup size 16}
\newcounter{k4}
\forloop{k4}{1}{\value{k4} < 21}{smallfriends-m-16-rep-\arabic{k4}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-16-rep-\arabic{k4}} \\ }

\section{Model $M_2$ versus $M_1$ lineup size 3}
\newcounter{k5}
\forloop{k5}{1}{\value{k5} < 21}{smallfriends-rev-m-3-rep-\arabic{k5}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-3-rep-\arabic{k5}} \\ }

\section{Model $M_2$ versus $M_1$ lineup size 6}
\newcounter{k6}
\forloop{k6}{1}{\value{k6} < 21}{smallfriends-rev-m-6-rep-\arabic{k6}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-6-rep-\arabic{k6}} \\ }

\section{Model $M_2$ versus $M_1$ lineup size 9}
\newcounter{k7}
\forloop{k7}{1}{\value{k7} < 21}{smallfriends-rev-m-9-rep-\arabic{k7}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-9-rep-\arabic{k7}} \\ }

\section{Model $M_2$ versus $M_1$ lineup size 12}
\newcounter{k8}
\forloop{k8}{1}{\value{k8} < 21}{smallfriends-rev-m-12-rep-\arabic{k8}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-12-rep-\arabic{k8}} \\ }

\section{Model $M_2$ versus $M_1$ lineup size 16}
\newcounter{k9}
\forloop{k9}{1}{\value{k9} < 21}{smallfriends-rev-m-16-rep-\arabic{k9}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-16-rep-\arabic{k9}} \\ }

\section{Model $M_1$ versus $M_3$ lineup size 3}
\newcounter{k10}
\forloop{k10}{1}{\value{k10} < 21}{smallfriends-eff2-m-3-rep-\arabic{k10}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-3-rep-\arabic{k10}} \\ }

\section{Model $M_1$ versus $M_3$ lineup size 6}
\newcounter{k11}
\forloop{k11}{1}{\value{k11} < 21}{smallfriends-eff2-m-6-rep-\arabic{k11}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-6-rep-\arabic{k11}} \\ }

\section{Model $M_1$ versus $M_3$ lineup size 9}
\newcounter{k12}
\forloop{k12}{1}{\value{k12} < 21}{smallfriends-eff2-m-9-rep-\arabic{k12}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-9-rep-\arabic{k12}} \\ }

\section{Model $M_1$ versus $M_3$ lineup size 12}
\newcounter{k13}
\forloop{k13}{1}{\value{k13} < 21}{smallfriends-eff2-m-12-rep-\arabic{k13}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-12-rep-\arabic{k13}} \\ }

\section{Model $M_1$ versus $M_3$ lineup size 16}
\newcounter{k14}
\forloop{k14}{1}{\value{k14} < 21}{smallfriends-eff2-m-16-rep-\arabic{k14}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-16-rep-\arabic{k14}} \\ }

\section{Model $M_3$ versus $M_1$ lineup size 3}
\newcounter{k15}
\forloop{k15}{1}{\value{k15} < 21}{smallfriends-eff2-rev-m-3-rep-\arabic{k15}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-3-rep-\arabic{k15}} \\ }

\section{Model $M_3$ versus $M_1$ lineup size 6}
\newcounter{k16}
\forloop{k16}{1}{\value{k16} < 21}{smallfriends-eff2-rev-m-6-rep-\arabic{k16}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-6-rep-\arabic{k16}} \\ }

\section{Model $M_3$ versus $M_1$ lineup size 9}
\newcounter{k17}
\forloop{k17}{1}{\value{k17} < 21}{smallfriends-eff2-rev-m-9-rep-\arabic{k17}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-9-rep-\arabic{k17}} \\ }

\section{Model $M_3$ versus $M_1$ lineup size 12}
\newcounter{k18}
\forloop{k18}{1}{\value{k18} < 21}{smallfriends-eff2-rev-m-12-rep-\arabic{k18}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-12-rep-\arabic{k18}} \\ }

\section{Model $M_3$ versus $M_1$ lineup size 16}
\newcounter{k19}
\forloop{k19}{1}{\value{k19} < 21}{smallfriends-eff2-rev-m-16-rep-\arabic{k19}\newline
\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-16-rep-\arabic{k19}} \\ }

\end{document}