### 2016-04-29 GGExperimentApr28/GGpresentationApr28.Rmd
add mini-experiment materials

---
title: "Network Lineups"
author: "Sam Tyner"
date: "April 28, 2016"
output: pdf_document
---

```{r, global_options, include=FALSE}
    library(knitr)
    library(formatR)
    opts_chunk$set(
      echo=FALSE,
      warning=FALSE,
      message=FALSE,
      fig.width = 8,
      fig.align = 'left', cache = FALSE)
```

## Intro

For each of the following lineups, write down the number of the panel that contains the graph you feel least resembles the graphs in the other panels.

For example, for the lineup below, you should pick 2.

```{r,echo = FALSE, fig.height=3}
datx <- data.frame(x = rnorm(30,2), y = rnorm(30,2), plot = 1)
daty <- data.frame(x = rnorm(30,2), y = rnorm(30,2), plot = 3)
datz <- data.frame(x = rnorm(30, 2), plot = 2)
datz$y = datz$x + rnorm(30, sd = .1)
dattt <- rbind(datx, daty, datz)
library(ggplot2)
ggplot(data = dattt) + geom_point(aes(x = x, y= y)) + theme(aspect.ratio = 1) + facet_wrap(~plot)
```

\newpage

```{r, echo=FALSE, warning=FALSE, message=FALSE}
setwd("~/Desktop/NetworksResearch/NetworksVizInference/GGExperimentApr28/")
for_gg <- data.frame(lineupname = rep(c("smallfriends", "smallfriends-rev", 
                                        "smallfriends-eff2", "smallfriends-eff2-rev"), each = 5) ,
                     M = rep(c(3,6,9,12,16), 4))
for_gg$rep <- c(17,6,8,4,16, 7,18,3,14,3, 10,5,7,11,4, 4,17,5,7,14)

for_gg$id <- 1:20

set.seed(138818)
for_gg$pres_order <- sample(LETTERS[1:20])

for_gg$filename <- paste0(for_gg$lineupname, "-m-", for_gg$M, "-rep-", for_gg$rep, ".csv")

# setwd("~/Desktop/NetworksResearch/NetworksVizInference/Data/lineupdata/")
# for (i in 1:nrow(for_gg)){
#   assign(x = paste0("dat",for_gg$pres_order[i]), value = read.csv(for_gg$filename[i]))  
# }
# for_gg$cat <- "CAT"
# for_gg$cat[c(1,5,9,13,17)] <- "noa_noc"
# for_gg$cat[(c(1,5,9,13,17) + 1)] <- "noa_c"
# for_gg$cat[(c(1,5,9,13,17) + 2)] <- "a_c"
# for_gg$cat[(c(1,5,9,13,17) + 3)] <- "a_noc"
# 
# all_dat <- NULL
# for(i in 1:nrow(for_gg)){
#   all_dat[[i]] <- read.csv(for_gg$filename[i])
# }
# 
# drink <- as.matrix(read.table("../s50_data/s50-alcohol.dat"))[,2]
# 
# dr_beh <- data.frame(friendid = 1:16, drink = drink[20:35])
# dr_beh$covariate <- "AH"
# dr_beh$covariate[which(dr_beh$drink %in% c(1,2))] <- "level A"
# dr_beh$covariate[which(dr_beh$drink %in% c(3,4))] <- "level B"

# library(ggplot2)
# library(geomnet)
# gg_plots <- NULL
# for (i in 1:length(all_dat)){
#     if (for_gg$cat[i] == "noa_noc"){
#       gg_plots[[i]] <- ggplot(data = all_dat[[i]], 
#                               aes(from_id = X1, to_id = X2)) +
#         geom_net(fiteach = T, size = 1, linewidth = .5) + 
#         facet_wrap(~plot_order) + theme_net() + 
#         theme(panel.background=element_rect(colour = 'black'))
#     } else if(for_gg$cat[i] == "a_noc"){
#       gg_plots[[i]] <- ggplot(data = all_dat[[i]], 
#                               aes(from_id = X1, to_id = X2)) +
#         geom_net(fiteach = T, size = 1, directed = T,  linewidth = .5, arrowsize = .1,
#                  arrowgap = .01) + 
#         facet_wrap(~plot_order) + theme_net() + 
#         theme(panel.background=element_rect(colour = 'black'))
#     } else if(for_gg$cat[i] == "noa_c"){
#       dat <- all_dat[[i]]
#       dat2 <- merge(dat, dr_beh, by.x = "X1", by.y = "friendid")
#       gg_plots[[i]] <- ggplot(data = dat2, aes(from_id = X1, to_id = X2)) + 
#         geom_net(fiteach = T, size = 1, linewidth = .5, aes(colour = covariate)) + 
#         facet_wrap(~plot_order) + theme_net() + 
#         theme(panel.background=element_rect(colour = 'black'))
#     } else{
#       dat <- all_dat[[i]]
#       dat2 <- merge(dat, dr_beh, by.x = "X1", by.y = "friendid")
#       gg_plots[[i]] <- ggplot(data = dat2, aes(from_id = X1, to_id = X2)) + 
#         geom_net(fiteach = T, size = 1, directed = T, arrowsize = .1, linewidth = .5,
#                  arrowgap = .01, aes(colour = covariate)) + 
#         facet_wrap(~plot_order) + theme_net() + 
#         theme(panel.background=element_rect(colour = 'black'))
#     }
# }
# savepdf <- function(file, width=16, height=10){
#   fname <- paste(file,".pdf",sep="")
#   pdf(fname, width=width/2.54, height=height/2.54,
#       pointsize=10)
#   par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.5,1.1,1))
# }
# for (i in 1:20){
#   savepdf(file = paste(for_gg$pres_order[i], sub(".csv", "", for_gg$filename[i])), 
#           width = 8*2.54, height = 8*2.54)
#   print(gg_plots[[i]])
#   dev.off()
# }
```

## A
![]("A smallfriends-eff2-rev-m-16-rep-14")
\newpage

## B
![]("B smallfriends-rev-m-6-rep-18")
\newpage

## C
![]("C smallfriends-eff2-m-3-rep-10")
\newpage

## D
![]("D smallfriends-eff2-rev-m-3-rep-4")
\newpage

## E
![]("E smallfriends-eff2-rev-m-6-rep-17")
\newpage

## F
![]("F smallfriends-rev-m-3-rep-7")
\newpage

## G
![]("G smallfriends-rev-m-16-rep-3")
\newpage

## H
![]("H smallfriends-eff2-m-16-rep-4")
\newpage

## I
![]("I smallfriends-rev-m-12-rep-14")
\newpage

## J
![]("J smallfriends-m-16-rep-16")
\newpage

## K
![]("K smallfriends-eff2-m-6-rep-5")
\newpage

## L
![]("L smallfriends-m-12-rep-4")
\newpage

## M
![]("M smallfriends-m-3-rep-17")
\newpage

## N
![]("N smallfriends-m-9-rep-8")
\newpage

## O
![]("O smallfriends-eff2-m-9-rep-7")
\newpage

## P
![]("P smallfriends-m-6-rep-6")
\newpage

## Q
![]("Q smallfriends-rev-m-9-rep-3")
\newpage

## R
![]("R smallfriends-eff2-rev-m-9-rep-5")
\newpage

## S
![]("S smallfriends-eff2-m-12-rep-11")
\newpage

## T
![]("T smallfriends-eff2-rev-m-12-rep-7")
\newpage