### 2016-07-10 JSM/VizInfSlides.Rmd
skeleton slides for JSM

---
title: "Visual Inference for Networks"
author: "Sam Tyner, Heike Hofmann"
date: "3 August 2016"
output: ioslides_presentation
---

## Outline

- Introduction
  * Visual Inference
  * Stochactic Actor-Oriented Models for Social Networks
- Simulation
- Experiment
- Results
- Future Work

# Introduction

## Visual Inference

## SAOMs

# Simulation

## RSiena

## ? SOMETHING ELSE ?

# Experiment

## Pilot Study (?)

## Procedure

# Results

## Summary

## Significance (?)

# Conclusion

## What we've learned

## Future Work