### 2016-07-12 JSM/VizInfSlides.Rmd
start slides

---
title: "Visual Inference for Networks"
author: "Sam Tyner, Heike Hofmann"
date: "3 August 2016"
output: ioslides_presentation
---

## Outline

- Introduction
  * Visual Inference
  * Stochactic Actor-Oriented Models for Social Networks
- Simulation
- Experiment
- Results
- Future Work

# Introduction

## Visual Inference

## Lineup (?)

## SAOMs

Goal: modeling dynamic networks using

- Network effects (e.g. outdegree, reciprocity)
- Actor covariate effects

## Model "Rules"

Observed networks treated as discrete observations of a continuous time Markov Chain.

1. At each (unobserved) time point in the CTMC one actor can change one of its ties.
2. Between network observations (say from $t_1$ to $t_2$) all actors are given the opportunity to change a tie at the same rate.
3. Once an actor can change, it tries to maximize its objective functon, $f_i(x)$
Parameters:

# Simulation

## RSiena

## ? SOMETHING ELSE ?

# Experiment

## Pilot Study (?)

## Procedure

# Results

## Summary

## Significance (?)

# Conclusion

## What we've learned

## Future Work

## References

## Thank you!