### 2016-07-12 JSM/VizInfSlides.Rmd
more sliding

---
title: "Visual Inference for Networks"
author: "Sam Tyner, Heike Hofmann"
date: "3 August 2016"
output: ioslides_presentation
---

## Outline

- Introduction
  * Visual Inference
  * Stochactic Actor-Oriented Models for Social Networks
- Simulation
- Experiment
- Results
- Future Work

# Introduction

## Visual Inference

## Lineups

$H_0$: "boring model" XX how specific? XX

$H_1$: "alternate model consisting of one additional statistically significant parameter" XX how specific? XX

Want: In a lineup, networks simulated from the $H_1$ model should appear so different from a series of models simulated from $H_0$ that an observer can pick out the different, alternative model.

## SAOMs

Goal: modeling dynamic networks using

- Network effects (e.g. outdegree, reciprocity)
- Actor covariate effects

## Model "Rules"

Observed networks treated as discrete observations of a continuous time Markov Chain.

1. At each (unobserved) time point in the CTMC one actor can change one of its ties.
2. Between network observations (say from $t_1$ to $t_2$) all actors are given the opportunity to change a tie at the same rate.
3. Once an actor can change, it tries to maximize its objective functon, $f_i(\mathbf{\beta},x) = \sum_k \beta_k s_{ki}(x)$, where $\beta_k$ are network or actor parameters, and $s_{ki}(x)$ are corresponding statistics.

# Simulation

Using a subset of the friendship network data from Michell & Amos we:

1. Fit three (XX two? XX) models 1000 times to obtain a distribution of parameter estimates
2. Simulated the second wave from each of our models to form lineups. Means from 1 were used as starting parameter values.
3. Estimated the parameters of each of our three (XX two? XX) models using the simulated second waves.

## RSiena

## ? SOMETHING ELSE ?

# Experiment

## Pilot Study (?)

## Procedure

# Results

## Summary

## Significance (?)

# Conclusion

## What we've learned

## Future Work

## References

## Thank you!