### 2017-09-23 ASApaper/paper.Rmd
add senate stuffs

---
title: Visual Inference for Goodness-of-Fit for A Network Model

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Sam Tyner
  thanks: The authors gratefully acknowledge ...
  affiliation: Department of Statistics, Iowa State University

- name: Heike Hofmann
  affiliation: Department of Statistics, Iowa State University

keywords:
- statistics
- social networks
- more

abstract: |
  The text of your abstract.  200 or fewer words.

bibliography: bibliography.bib
output: rticles::asa_article
---

# Introduction

# Data

# Model Definition and Choice

## Choosing appropriate models

In order to determine which models to explore for application of our novel goodness-of-fit testing method, we first fit many possible models and selected a few significant effects. To determine the significance of the effects, we followed this procedure:

1. Define the simple effects structure of the data: the rate parameters and the outdegree and reciprocity parameters. 
2. Add each possible effect one-at-a-time to the model structure, as determined by the effects documentation function.
3. Fit each model to the data and check for convergence.
    a. If the model converged, move to 4.
    b. If the model did not converge, use the previous fitted values as starting values and repeat 5 times or until convergence, whichever comes first.
4. Test the added parameter for significance using the Wald-type test.
5. Report out the estimate of the additional parameter, its standard error, Wald $p$ value, and convergence criterion.

After completing the procedure for all model effects, we selected effects to include that converged and had a Wald $p$-value of less than 0.10, and seemed to have a reasonable interpretation for our data.

# Simulation Procedure

1. Fit each model 1,000 times and save the parameter estimates, standard errors, and overall convergence criterion.
2. (From the models that converged ?) Compute the mean of the 1,000 estimates of each parameter.
3. Use the mean values from step 2 as the parameter values in the model and simulate 100 observations from the model. These simulated networks will become the "true data."
4. To each simulation, refit the model form from which it was simulated. 
5. Repeat steps 4-5 for various permutations of model forms and parameter values; e.g. double the parameter values, quadruple the parameter values, change the signs on the parameters, etc.
6. Simulate additional "alternative data" from the model in step 3 and the fitted model in step 4. 
7. Place the appropriate node-link diagram of each "true data" observation in a lineup with $M-1$ (M=12? M=9?) "alternative data" node-link diagrams. 
8. Show the lineups to participants, with no participant seeing the same "true data" or "alternative data" plot twice.

<!--
This template demonstrates some of the basic latex you'll need to know to create a ASA article.

\section{Verifications}
\label{sec:verify}
This section will be just long enough to illustrate what a full page of
text looks like, for margins and spacing.

@Campbell02 @Schubert13 [@Chi81]

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
__With this spacing we have 30 lines per page.__
-->