### 2017-09-27 ASApaper/paper.Rmd
discussion in paper

---
title: Visual Inference for Goodness-of-Fit for A Network Model

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Sam Tyner
  thanks: The authors gratefully acknowledge ...
  affiliation: Department of Statistics, Iowa State University

- name: Heike Hofmann
  affiliation: Department of Statistics, Iowa State University

keywords:
- statistics
- social networks
- more

abstract: |
  The text of your abstract.  200 or fewer words.

bibliography: bibliography.bib
output: rticles::asa_article
---

```{r pkgs, echo = FALSE}
library(tidyverse)
library(RSienaTest)
library(RSiena)
library(geomnet)
library(GGally)
```

# Introduction

# Data

# Model Definition and Choice

## Choosing appropriate models

In order to determine which models to explore for application of our novel goodness-of-fit testing method, we first fit many possible models and selected a few significant effects. To determine the significance of the effects, we followed this procedure:

1. Define the simple effects structure of the data: the rate parameters and the outdegree and reciprocity parameters. 
2. Add each possible effect one-at-a-time to the model structure, as determined by the effects documentation function.
3. Fit each model to the data and check for convergence.
    a. If the model converged, move to 4.
    b. If the model did not converge, use the previous fitted values as starting values and repeat 5 times or until convergence, whichever comes first.
4. Test the added parameter for significance using the Wald-type test.
5. Report out the estimate of the additional parameter, its standard error, Wald $p$ value, and convergence criterion.

After completing the procedure for all model effects, we selected effects to include that converged and had a Wald $p$-value of less than 0.10, and seemed to have a reasonable interpretation for our data.

The most significant effect was the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -6 with a standard error of 0.11. This effect relies on the number of transitive closures formed between two senators from different parties. Due to the divisive nature of American politics, the party effect is dominant. Another significant effect was the same JTT parameter for the sex covariate, with an estimate of about 3 with a standard error of 0.89. Another significant effect was the covariate-related similarity score-weighted transitive triplets for the number of bills authored by a senator. This effect was estimated at about 10 with standard error of 3.9, and the high positive effect suggests senators tend to collaborate with other senators who author about the same number of bills they do.

# Simulation Procedure

1. Fit each model 1,000 times and save the parameter estimates, standard errors, and overall convergence criterion.
2. (From the models that converged ?) Compute the mean of the 1,000 estimates of each parameter.
3. Use the mean values from step 2 as the parameter values in the model and simulate 100 observations from the model. These simulated networks will become the "true data."
4. To each simulation, refit the model form from which it was simulated. 
5. Repeat steps 4-5 for various permutations of model forms and parameter values; e.g. double the parameter values, quadruple the parameter values, change the signs on the parameters, etc.
6. Simulate additional "alternative data" from the model in step 3 and the fitted model in step 4. 
7. Place the appropriate node-link diagram of each "true data" observation in a lineup with $M-1$ (M=12? M=9?) "alternative data" node-link diagrams. 
8. Show the lineups to participants, with no participant seeing the same "true data" or "alternative data" plot twice.

# Exploratory plots

```{r M1}
M1senateEsts <- read_csv("../Data/senate/basicModelFitsSenate1000.csv")
inisdSenateEsts <- read_csv("../Data/senate/inisdModelFitsSenate1000.csv")
jttpSenateEsts <- read_csv("../Data/senate/jttpModelFitsSenateAll1000.csv")
jttsSenateEsts <- read_csv("../Data/senate/jttsModelFitsSenateAll1000.csv")
# ggpairs(M1senateEsts, columns = 1:8)
# M1senateEsts %>% mutate(converged = (maxConv <= 0.25)) %>% 
#   ggplot() + 
#   geom_density(aes(x = density, color = converged), adjust = .5) 
# M1senateEsts %>% mutate(converged = (maxConv <= 0.25)) %>% 
#   ggplot() + 
#   geom_histogram(aes(x = density, color = converged), binwidth = .005, fill = NA) + facet_wrap(~converged)

# make ests long - M1
M1senateEsts %>% 
  select(Rate:recip, maxConv, sim) %>% 
  gather(parameter, estimate, Rate:recip) -> melt1 
M1senateEsts %>% 
  select(se_density:sim) %>% 
  gather(parameter, se, se_density:se_recip) %>% 
  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
left_join(melt1, melt2) %>% arrange(sim) -> M1SenEstsLong

# make ests long - inisd
inisdSenateEsts %>% 
  select(Rate:inIsDegree, maxConv, sim) %>% 
  gather(parameter, estimate, Rate:inIsDegree) -> melt1 
inisdSenateEsts %>% 
  select(se_density:sim) %>% 
  gather(parameter, se, se_density:se_inIsDegree) %>% 
  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
left_join(melt1, melt2) %>% arrange(sim) -> inisdSenEstsLong

# make ests long - jttp
jttpSenateEsts %>% 
  select(Rate:jumpXTransTrip, maxConv, sim) %>% 
  gather(parameter, estimate, Rate:jumpXTransTrip) -> melt1 
jttpSenateEsts %>% 
  select(se_density:maxConv, sim) %>% 
  gather(parameter, se, se_density:se_jumpXTransTrip) %>% 
  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
left_join(melt1, melt2) %>% arrange(sim) -> jttpSenLong

# make ests long - jtts
jttsSenateEsts %>% 
  select(Rate:jumpXTransTrip, maxConv, sim) %>% 
  gather(parameter, estimate, Rate:jumpXTransTrip) -> melt1 
jttsSenateEsts %>% 
  select(se_density:maxConv, sim) %>% 
  gather(parameter, se, se_density:se_jumpXTransTrip) %>% 
  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
left_join(melt1, melt2) %>% arrange(sim) -> jttsSenLong

```

<!--
This template demonstrates some of the basic latex you'll need to know to create a ASA article.

\section{Verifications}
\label{sec:verify}
This section will be just long enough to illustrate what a full page of
text looks like, for margins and spacing.

@Campbell02 @Schubert13 [@Chi81]

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
__With this spacing we have 30 lines per page.__
-->