### 2017-10-11 GG-10-12/slides.Rmd
slides

---
title: "Applying visual inference to a social network model"
author: "Sam Tyner"
date: "10/12/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = '100%', message = FALSE, warning = FALSE,
                      fig.align = 'center')
```

```{r}
library(tidyverse)
library(RSiena)
library(geomnet)
```

## Training 1

*Which plot is most different from the others?*

```{r train1}
dat <- read_csv("LineupData/samepx1sd234567pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = 'fruchtermanreingold') + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Training 2

*Which plot is most different from the others?*

```{r train2}
dat <- read_csv("LineupData/jttsx5sd234567picknull.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = 'fruchtermanreingold') + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial A
<!--basic 1.3-->
```{r trialA}
dat <- read_csv("LineupData/basicx1.3sd123456pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = 'fruchtermanreingold') + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial B
<!--jttp -1.5-->

```{r trialB}
dat <- read_csv("LineupData/jttpx-1.5sd98765pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial C
<!--jtts 2-->

## Trial D
<!--simttb 2-->

## Trial E
<!--basic-recip -5-->

## Trial F
<!--basic -1-->

## Trial G
<!--samettp 2-->

## Trial H
<!--basic-density -1-->

## Trial I
<!--basic-recip 1.5-->

## Trial J
<!--jtts -5-->

## Trial K
<!--samettp 2-->

## Trial L
<!--simttb -1-->

## Trial M
<!--jttp 4-->

## Trial N
<!--basic-density 1.2-->

## Trial O
<!--jttp 4-->

## Trial P
<!--samettp -1 -->