### 2017-10-12 GG-10-12/slides.Rmd

---
title: "Applying visual inference to a social network model"
author: "Sam Tyner"
date: "10/12/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = '100%', message = FALSE, warning = FALSE,
                      fig.align = 'center')
```

```{r}
library(tidyverse)
library(RSiena)
library(geomnet)
library(igraph)
```

## Training 1

*Which plot is most different from the others?*

```{r train1}
dat <- read_csv("LineupData/samepx1sd234567pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = 'fruchtermanreingold') + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Training 2

*Which plot is most different from the others?*

```{r train2}
dat <- read_csv("LineupData/jttsx5sd234567picknull.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = 'fruchtermanreingold') + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial A
<!--basic 1.3-->
```{r trialA}
dat <- read_csv("LineupData/basicx1.3sd123456pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = 'fruchtermanreingold') + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial B
<!--jttp -1.5-->

```{r trialB}
dat <- read_csv("LineupData/jttpx-1.5sd98765pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial C
<!--jtts 2-->

```{r trialC}
dat <- read_csv("LineupData/jttsx2sd123456pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial D
<!--simttb 2-->

```{r trialD}
dat <- read_csv("LineupData/simttbx2sd123456pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial D2

```{r trialD2}
#same as K with KK instead of FR layout
dat <- read_csv("LineupData/samettpx-1sd35792pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial E
<!--basic-recip -5-->

```{r trialE}
dat <- read_csv("LineupData/basicx-5sd346231pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5,
                 layout.alg = 'fruchtermanreingold') + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial F
<!--basic -1-->

```{r trialF}
dat <- read_csv("LineupData/basicx-1sd346231pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = "fruchtermanreingold") + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial G
<!--samettp 2-->

```{r trialG}
dat <- read_csv("LineupData/samettpx2sd23452pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial H
<!--basic-density -1-->

```{r trialH}
dat <- read_csv("LineupData/basicx-1sd987234pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = "fruchtermanreingold") + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial I
<!--basic-recip 1.5-->

```{r trialI}
dat <- read_csv("LineupData/basicx1.5sd232342pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = "fruchtermanreingold") + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial J
<!--jtts -5-->

```{r trialJ}
dat <- read_csv("LineupData/jttsx-5sd45745pickalt.csv")

ggplot(data = dat, aes(from_id=from, to_id=to, colour=log(csize))) + 
        geom_net(arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 arrowgap = .015, arrowsize = .3, size = .75, layout.alg = 'fruchtermanreingold') + 
        theme_net() +
        theme(panel.background = element_rect(color = 'black')) +
        scale_colour_gradient(low="#f768a1", high = "#49006a") + facet_wrap(~ord) +
        theme(legend.position="none")
```

## Trial K
<!--samettp -1-->

```{r trialK}
dat <- read_csv("LineupData/samettpx-1sd35792pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = 'fruchtermanreingold') + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial L
<!--simttb -1-->

```{r trialL}
dat <- read_csv("LineupData/simttbx-1sd234624pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial M
<!--jttp 4-->

```{r trialM}
dat <- read_csv("LineupData/jttpx4sd234624pickalt.csv")
df <- data.frame(expand.grid(
        from=paste0("V", 1:155), to = paste0("V", 1:155)), stringsAsFactors = FALSE)
netnest <- dat %>% group_by(sim) %>% nest()
netnest <- netnest %>% mutate(
        data = data %>% purrr::map(.f = function(d) {
          adj1 <- d %>% right_join(df) %>% 
            xtabs(!is.na(dep.var.id)~from+to, data =.) %>% as.matrix()

          ig1 <- graph_from_adjacency_matrix(adj1)
          comps <- components(ig1)
          clu_df <- data.frame(from=names(comps$membership), clu = comps$membership)
          clu_df$csize=comps$csize[clu_df$clu]
          d %>% left_join(clu_df, by="from")
        })
) 
dat <- unnest(netnest, data)
ggplot(data = dat, aes(from_id=from, to_id=to, colour=log(csize))) + 
        geom_net(arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 arrowgap = .015, arrowsize = .3, size = .75, layout.alg = 'fruchtermanreingold') + 
        theme_net() +
        theme(panel.background = element_rect(color = 'black')) +
        scale_colour_gradient(low="#f768a1", high = "#49006a") + facet_wrap(~ord) +
        theme(legend.position="none")
```

## Trial N
<!--basic-density 1.2-->

```{r trialN}
dat <- read_csv("LineupData/basicx1.2sd34231pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```