### 2017-10-19 ASApaper/paper.Rnw
introduction about 2/3 ish done

\documentclass[12pt]{article}
% \usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
% \geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\usepackage{graphicx}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{animate}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
% \usepackage{color}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[dvipsnames,svgnames,table]{xcolor}
\usepackage{hyperref}
\graphicspath{{figure/}}
% \usepackage{endfloat} % Figures to the end of the document

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
%---------------------------------------------------
%                 Editing Commands
%\newcommand{\hh}[1]{{\color{magenta} #1}}
%\newcommand{\st}[1]{{\color{orange} #1}}

%---------------------------------------------------
%                 Placing Figures
\renewcommand{\topfraction}{0.99}	% max fraction of floats at top
\renewcommand{\bottomfraction}{0.8}	% max fraction of floats at bottom
\renewcommand{\textfraction}{0.05}	% allow minimal text w. figs

%---------------------------------------------------
% Define new environment
\newtheorem{theorem}{Theorem}[section]
\newtheorem{algorithm}[theorem]{Algorithm}
%---------------------------------------------------

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Visual Inference for Significance and Goodness-of-Fit Testing for a Social Network Model}
  \author{Samantha Tyner\thanks{
    The authors gratefully acknowledge funding from the National Science Foundation Grant \# DMS 1007697. All data collection has been conducted with approval from the Institutional Review Board IRB 10-347}\hspace{.2cm}\\
    Department of Statistics and Statistical Laboratory, Iowa State University\\
    and \\
    Heike Hofmann\\
    Department of Statistics and Statistical Laboratory, Iowa State University}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Visual Inference for Significance and Goodness-of-Fit Testing for a Social Network Model}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Abstract goes here.
\end{abstract}

\noindent%
{\it Keywords:} social network analysis, visual inference, dynamic networks, network visualization, network mapping, goodness-of-fit, hypothesis testing
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\tableofcontents
\newpage

<<child-demo, child='paper-child.Rnw'>>=
@

\bibliographystyle{asa}
\bibliography{bibliography}

\end{document}