### 2019-03-05 ASApaper/paper.Rnw

\documentclass[12pt]{article}
% \usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
% \geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\usepackage{graphicx}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
%\usepackage{animate}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
%\usepackage{color}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[dvipsnames,svgnames,table]{xcolor}
\usepackage{hyperref}
\usepackage{numprint}

\graphicspath{{figure/}}
% \usepackage{endfloat} % Figures to the end of the document
\newcommand{\st}[1]{{\color{orange} #1}}
\newcommand{\hh}[1]{{\color{magenta} #1}}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
%---------------------------------------------------
%                 Editing Commands
%\newcommand{\hh}[1]{{\color{magenta} #1}}
%\newcommand{\st}[1]{{\color{orange} #1}}

%---------------------------------------------------
%                 Placing Figures
\renewcommand{\topfraction}{0.99}	% max fraction of floats at top
\renewcommand{\bottomfraction}{0.8}	% max fraction of floats at bottom
\renewcommand{\textfraction}{0.05}	% allow minimal text w. figs

%---------------------------------------------------
% Define new environment
\newtheorem{theorem}{Theorem}[section]
\newtheorem{algorithm}[theorem]{Algorithm}
%---------------------------------------------------

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Visual Inference for a Social Network Model}
  \author{Samantha Tyner\thanks{
    The authors gratefully acknowledge funding from the National Science Foundation Grant \# DMS 1007697. All data collection has been conducted with approval from the Institutional Review Board IRB 10-347}\hspace{.2cm}\\
    Department of Statistics and Statistical Laboratory, Iowa State University\\
    and \\
    Heike Hofmann\\
    Department of Statistics and Statistical Laboratory, Iowa State University}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Visual Inference for a Social Network Model}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Three of the most important assessments of a statistical model are significance tests of parameters, goodness-of-fit tests, and power calculations of the tests. All three tasks become more difficult as the model becomes more complex. We will explore these three assessments of one particularly complex set of models, continuous time Markov chain (CTMC) models, for dynamic social networks  \citep{saompaper}. In this paper, we propose new methods for significance and goodness-of-fit testing, as well as power calculations for CTMC models via the visual inference (VI) paradigm of \citet{Bujaetal}. With VI, we can look at entire datasets simulated from a model, instead of relying a single metric such as a $p$-value. We conducted a VI experiment, with participants recruited via Amazon Mechanical Turk, to assess the significance of CTMC parameters, the fit of a CTMC model, and the visual power of CTMC parameters.  
\end{abstract}

\noindent%
{\it Keywords:} social network analysis, visual inference, dynamic networks, network visualization, network mapping, goodness-of-fit, hypothesis testing
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

%\tableofcontents
\newpage

<<child-demo, child='paper-child.Rnw'>>=
@

\bibliographystyle{asa}
\bibliography{bibliography}

\end{document}